### 2023-01-22 definitions.qmd
start project

# Definitions {.unnumbered}

Sequestration: Process of isolating and storing respired CO<sub>2</sub> within the ocean interior and away from atmospheric exchanges

Export flux:

respiration

photosynthesis

BCP strength

BCP efficiency

Net Primary Production: Net conversion rate of inorganic carbon to organic carbon by phytoplankton photosynthesis [@Siegel2022-wd]

Euphotic Zone: Depth range in the near-surface ocean with enough sunlight for phytoplankton net primary production to occur [@Siegel2022-wd]

Twiligth Zone = Mesopélagic zone

Mixed Layer

Remineralization: Conversion of organic matter back to inorganic forms by heterotrophic microbes and animals [@Siegel2022-wd]

Particle size distribution (or particle size spectrum)

Carbon dioxide removal