### 2023-02-15 annexes.qmd
update

# Annexes

```{=html}
<!--#https://github.com/quarto-dev/quarto-cli/discussions/1042-->
<style type="text/css">
caption, .table-caption {
  text-align: left;
}
</style>
```

```{r, poc database from traps}
#| label: tbl-poc-from-sediment-traps
#| tbl-cap: Compilation of POC flux measurements from sediment traps. Data taken from public databases listed on PANGAEA or extracted from articles (with the help of Frédéric Le Moigne and Meike Vogt). The number of measurements could be different from the original publication after data cleaning and potential removal of duplicates.
#| tbl-cap-location: bottom

library(dplyr)
library(knitr)
library(kableExtra)
library(purrr)

# read data
trap <- vroom::vroom('DATA/Trap_Thorium/TRAP_DATABASE.csv')
#thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')

# clean a bit
trap[trap$doi == 'https://doi.org/10.1007/s00227-018-3376-1',]$doi <- '10.1007/s00227-018-3376-1'
trap[trap$doi == 'https://doi.org/10.1016/j.dsr.2016.08.015',]$doi <- '10.1016/j.dsr.2016.08.015'
trap[trap$doi == 'Geochimica et Cosmochimica Acta, Vol. 66, No. 3, pp. 457–465, 2002',]$doi <- '10.1016/S0016-7037(01)00787-6'

# some stats
trap_resume <- trap |> group_by(doi) |> dplyr::summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))

# BASED ON : https://stackoverflow.com/questions/48506712/kable-kableextra-cells-with-hyperlinks
# make url links
make_link <- function(doi){
  link <- paste0('https://doi.org/',doi)
  return(link)
}
url <- map_chr(trap_resume$doi, make_link)

# # put url in dataframe
# trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, 'html', link = url)) |> select(doi_url, data_points) |>
#   kable('html', escape = FALSE, col.names = c('Sources of data', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))

trap_resume <- trap_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l')

# make final table
trap_resume
```

```{r, poc database from thorium}
#| label: tbl-poc-from-thorium
#| tbl-cap: Compilation of POC flux measurements derived from Thorium 234. Data taken from public databases listed on PANGAEA or extracted from articles (with the help of Frédéric Le Moigne). The number of measurements could be different from the original publication after data cleaning and potential removal of duplicates.
#| tbl-cap-location: bottom

# read data
thorium <- vroom::vroom('DATA/Trap_Thorium/THORIUM_DATABASE.csv')

# some stats
thorium_resume <- thorium |> group_by(doi) |> summarise(data_points = n(), max_date = lubridate::year(max(start_date)), min_date = lubridate::year(min(start_date))) |> arrange(desc(data_points)) |>
  mutate(period_covered = if_else(max_date > min_date, paste0(min_date,'-',max_date), paste0(max_date)))
thorium_resume$doi <- c('10.1594/PANGAEA.809717', '10.1029/2018GB00615', '10.1525/elementa.2020.030')

# make url links
url <- c('https://doi.org/10.1594/PANGAEA.809717', 'https://doi.org/10.1029/2018GB006158', 'https://doi.org/10.1525/elementa.2020.030')

# put url in dataframe
thorium_resume <- thorium_resume |> mutate(doi_url = cell_spec(doi, link = url)) |> select(doi_url, period_covered, data_points) |>
  kable(escape = FALSE, col.names = c('Source of data', 'Period covered', 'Number of measurements'), align = 'l') #|> kable_styling(bootstrap_options = c("hover", "condensed"))

# make final table
thorium_resume
```