### 2023-03-16 bcp_sequestration.qmd
update

# Revisiting the BCP with the CONVERSE framework

add a short description and then write below a sumbimtted dradft to Nature Geoscience (currently in revision)

```{=latex}
\includepdf[pages=-]{"/home/flo/dox/THESIS/Ricour2023main"} %https://tex.stackexchange.com/questions/183469/includepdf-only-shows-first-page thank you stranger
\includepdf[pages=-]{"/home/flo/dox/THESIS/Ricour2023supp"}
```