### 2023-07-26 before-body.tex
clean repo

\documentclass[a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{titlesec}
\usepackage{geometry} % Added for adjusting the page margins

\titleformat{\chapter}[display]
{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{20pt}{\Huge}

% Added for adjusting the page margins
\geometry{
    left=3cm,
    right=3cm,
    top=1cm,
    bottom=3cm
}

\begin{document}

\begin{titlepage}
    \begin{center}
        \includegraphics[width=0.3\textwidth]{FRS-FNR_S2-PANT_UK_CS.png} % Insert your university logo here
        \vspace{1cm}

        \LARGE \textbf{University of Liège}\\
        \LARGE \textbf{Sorbonne University}\\

        \vspace{1.5cm}
        %{\huge \textbf{Doctoral Thesis}}\\[0.5cm]

        \hrulefill \\
        \vspace{0.5cm}
        {\huge \textbf{Towards a new insight of the carbon transport in the global ocean}} \\
        \hrulefill

        \vspace{1.5cm}
        {\Large \textbf{Florian Ricour}}

        \vspace{1cm}

        Dissertation presented in fulfillment for the degree of Doctor of Sciences

        \vfill

        \begin{flushleft}
            \large
            \textbf{Supervisors}\\
            \vspace{0.1cm}
            Hervé Claustre, director\\
            Marilaure Grégoire, director\\
            Alexander Barth, co-director\\
        \end{flushleft}

        \begin{flushleft}
            \large
            \textbf{Jury Members}\\
            \vspace{0.1cm}
            \begin{tabular}{lll}
                Pr. Stephanie Henson & University of Southampton & Reviewer\\
                Dr. Margaret Estapa & University of Maine & Reviewer \\
                Dr. Griet Neukermans & University of Ghent & Examiner \\
                Dr. Ivona Cetini\'{c} & Morgan State University & Examiner\\
                Pr. Lars Stemmann & Sorbonne University & Examiner\\
                Dr. Lionel Guidi & Sorbonne University & Guest\\
                Dr. Hervé Claustre & Sorbonne University & PhD Director\\
                Pr. Marilaure Grégoire & University of Liège & PhD Director\\
                Dr. Alexander Barth & University of Liège & PhD Co-director\\
            \end{tabular}
        \end{flushleft}

        \vspace{1cm}
        {\large May 2023} % Replace with the month and year of submission

    \end{center}
\end{titlepage}

\newpage
\thispagestyle{empty}
\null

\end{document}